### 2013-08-15 presentation/presentation.tex
Illustrated and fixed hierarchical theorem

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\em i.e.},~}
\newcommand{\eg}{{\em e.g.},~}
\newcommand{\parentd}{../}

\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage

\title{Matroids from families of null hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}

\maketitle

\begin{frame}
\frametitle{Outline}
\setcounter{tocdepth}{3}
\tableofcontents
\end{frame}

\section{Logically dependent hypotheses}

\begin{frame}
\frametitle{Pairwise comparisons, some history }

Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
\begin{itemize}
  \item Parameters: $\mu_1,\mu_2,\mu_3$ ; \\ Hypotheses: $H_1:\mu_1=\mu_2,H_2:\mu_2=\mu_3;H_3:\mu_1=\mu_3$
  \item A level-$\alpha$ test of $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 : \mu_1=\mu_2=\mu_3$. followed by level-$\alpha$ tests of $H_1,H_2,H_3$ protects the FWER at level $\alpha$.
  \begin{itemize}
    \item  essentially closed test; no need to test pairwise intersection hypotheses since $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j$,
    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise comparisons, some history}

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
       \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Groups}
\input{full4ext}
\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{itemize}

\item Consider $\mu_i, i \in \{1,...,m\}$

\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Means}
\input{full4}
\end{frame}

\begin{frame}
\frametitle{General contrast tests (two-sided)}
\begin{itemize}

\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$

\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
  c_{ij} = 0$
\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
    I}$
\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$

\item Elementary hypotheses are represented by column vectors,
  intersection hypotheses by sub-matrices of the contrast matrix

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 2: Contrast test}

\begin{tabular}{ll}
  \begin{minipage}{.45\textwidth}
    All pairwise comparisons of 4 means 
    \begin{displaymath}
    \bs{C}^t = \left(
      \begin{array}{llll}
      -1 & 1 & 0 & 0 \\
      -1 & 0 & 1 & 0 \\
      -1 & 0 & 0 & 1 \\
      0 & -1 & 1 & 0 \\
      0 & -1 & 0 & 1 \\
      0 & 0 & -1 & 1 
    \end{array}
    \right)
    \end{displaymath}
  \end{minipage} &
  \begin{minipage}{.55\textwidth}
    Modified Williams Contrasts for Dose Response
    \begin{displaymath}
      \bs{C}^t=\left(
      \begin{array}{llll}
        1 & 0 & 0 & -1\\
        1 & 0 &-.51 & -.49 \\
        1 & -.35 & -.33 & -.32\\
        .51 & .49 & -.51 & -.48\\
        .51 & .49 & 0 &  -1\\
        .35 & .33 & .32& -1
      \end{array}\right)
    \end{displaymath}
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependent null hypotheses}

\begin{itemize}
\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
  dependence relation $\sim_H$ on $\mathcal{P}(S)$
\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq  H_{I\setminus i}$
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
  column rank
\end{itemize} % ends low level
\end{frame}

\section{Matroids}

\begin{frame}
\frametitle{Matroids}
\begin{itemize}

\item A matroid on $S$ endows $\mathcal{P}(S)$ with an algebraic
  structure 
\item Matroids generalize the structures of graphs and vector spaces,
  \eg circuits, bases, independent sets
\item Many ways to define a matroid
  \begin{itemize}
  \item From a graph perspective: Circuit axioms
  \item From a vector space perspective: Independence axioms
  \end{itemize}
\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
\item \textbf{/Some illustration hereabouts/}
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Circuit axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{cycle}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
    \begin{enumerate}
    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
      $\exists C_3 \subseteq (C_1 \cup C_2)$,
    \end{enumerate}
    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
  \end{minipage}
\end{tabular}

\end{frame}

% \item \textbf{/Some illustration hereabouts/}

% \end{itemize} % ends low level
% \end{frame}

\begin{frame}
\frametitle{Independence axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
  \only<1>{\input{independent}}
  \only<2>{\input{cycle}}
  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$.

  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependence axioms}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{dependent}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  Relation $\sim_M$ on elements and subsets of $S$,
  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
  \begin{enumerate}
  \item $y \sim Y$ for all $y \in Y$
  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
    then $y \sim x \cup (Y \setminus y)$
  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
  \end{enumerate}
  then $\sim_M$ defines a matroid on $S$.
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}{Maximally independent sets}
  \begin{itemize}
  \item Bases are maximally independent sets of $S$
  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
  \item Circuits are minimally dependent sets, \ie each subset of $C$
    with rank $\rho C - 1$ is a maximally independent set of $C$
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Maximally dependent sets}
\begin{itemize}
\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
\item Closed sets {\em flats}: $\mathscr{L} = \{I \subseteq S: \sigma I = I\}$
\item Hyperplanes are flats of rank $\rho S - 1$
\item All flats are given by closing hyperplanes under intersection
\item Circuits in $M^*$ are the complements of hyperplanes in $M$
\end{itemize} % ends low level
\end{frame}

\begin{frame}{Examples}
\framesubtitle<1>{Independent set}  
\framesubtitle<2>{Base}  
\framesubtitle<4>{Dependent set}  
\framesubtitle<3>{Cycle}  
\framesubtitle<5>{Closure}  
\only<1>{\input{independent}}
\only<2>{\input{base}}
\only<4>{\input{dependent}}
\only<3>{\input{cycle}}
\only<5>{\input{closure}}
\end{frame}

\section{Matroids from logically dependent hypotheses}

\begin{frame}
\frametitle{Matroids from logically dependent hypotheses}

  \begin{block}{Theorem}
    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
    S$, an intersection hypotheses $H_I$ is independent iff the
    corresponding edge set $\{e_i:i \in I\}$ is independent in the
    cycle matroid on $\mathscr{G}_I$.

For a family of two-sided
    general contrast tests $H_i$, $i \in S$, an intersection
    hypotheses $H_I$ is independent iff the corresponding contrast
    matrix $\bs{C}_I$ is independent in the vectorial matroid on
    $\mathbb{R}^m$.

    \begin{displaymath}
      i \sim_H I \Leftrightarrow i \sim_M I
    \end{displaymath}
  \end{block}

\end{frame}

\begin{frame}
\frametitle{Closed test procedure}
\begin{itemize}

\item The flats of a matroid are closed under intersection

\item Every elementary hypothesis is a flat

\item Using local level-$\alpha$ tests for each flat we have a closed test procedure that controls the FWER in the strong sense (Closed test principle)

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Examples}
\begin{itemize}

\item Pairwise comparisons

\item The flats of the complete graph correspond to the partitions of
  $\{1,...,m\}$
\item This provides efficient methods to generate all flats for
  pairwise mean comparisons \cite{er1988fast,knuth2005art,kokosinski2006new}

\item The flats of a matroid generate a lattice, procedures testing not all possible comparisons correspond to intervals on that lattice

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Connections to existing procedures}

\textbf{Old wine in new bottles ((c) L. Hothorn)}
\begin{itemize}
\item \cite{westfall1997multiple} checks each intersection hypotheses
  whether it is flat using dependence arguments.
\item \cite{westfall2007multiple} provide a branch and bound algorithm
  to generate all flats. They note that for pairwise comparisons all
  exhaustive hypotheses are {\em partition hypotheses} 
\item \cite{bergmann1988improvements} define the concepts of an
  exhaustive sets
\item \cite{bernhard1991computergestuetzte} shows properties of
  exhaustive sets that are basically {\em closure axioms} of a matroid
  and provides algorithms to check each intersection hypotheses if
  flat using rank function.

  \cite{hommel1999bonferroni} show {\em closure axioms} for exhaustive
  sets 
\item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)

\item Lemma 1.5 in (Bernhard) about weakly exhaustive sets are ``closure axioms'' of a matroid

\item Graphical procedures for pairwise comparisons (Weichert) identify dependence using cycle in the graph

\item \emph{However:}
\begin{itemize}
\item proofs are shorter (more elegant),
\item efficient algorithms exist to find the flats of matroids,
\item we may define procedures using independent sets,
\item \ldots{}
\end{itemize}

\end{itemize} % ends low level
\end{frame}

\begin{frame}{Matroid procedures for pairwise mean comparisons}

\begin{tabular}{ll}
  \begin{minipage}{.3\textwidth}
    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
     {\tiny wikipedia:Bell\_number}
  \end{minipage} &
  \begin{minipage}{.7\textwidth}
  \begin{itemize}
  \item Two-sided comparisons between $m$ means
  \item The cycle matroid of the corresponding graph representation
    provides an equivalent dependence structure
  \item The flats of a (simple) complete graph are
    isomporphic to the partitions of a set of size $m$
  \item Provides an efficient way to generate all exhaustive sets by
    enumeration of all partitions or hyperplanes \cite{er1988fast,knuth2005art,kokosinski2006new} 
  \item The number of flats is given by the Bell-numbers \ie 5, 15,
    52, 203, 877, 4140, ... 
  \end{itemize}
  \end{minipage}
\end{tabular}

\end{frame}

\section{Graphical approaches for logically dependent hypotheses}

\begin{frame}
\frametitle{Reminder: Graphical approaches}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
    \end{tikzpicture}

  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item Provides weights $w_{i,I}$ for each elementary hypotheses
      $H_i$ and each non-empty subset $I \subset S$
    \item Weights are used to define a closed test procedure of
      weighted intersection hypotheses tests
    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
    \end{itemize}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Example: Hierarchical LSD}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \input{lsd2.tex}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{itemize}
    \item 3 Hypotheses:
      \begin{enumerate}
      \item Placebo vs. High Dose
      \item Placebo vs. Low Dose
      \item High Dose vs. Low Dose
      \end{enumerate}
    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
    \item Test $H_1$ at level $\alpha$
    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
      simultaneously
    \item Can we generalize this?
    \end{itemize}
  \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
  \frametitle{Fixed sequence test of pairwise mean comparisons}
\begin{itemize}
\item {\bf Picture of hierarchical test hereabouts}
\end{itemize}

\begin{block}{Theorem}
  Assume that $M(\mathscr{H})$ neither contains loops nor
  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are to
  be tested using at level $\alpha_r$ according to a prefixed order.
  If for some $1 \leq s \leq t$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha_ra$.
\end{block}

\end{frame}

\begin{frame}
  \frametitle{Example: Multiple doses}

\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
    \begin{tikzpicture}[overlay]
      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
    \end{tikzpicture}
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{tikzpicture}[remember picture,overlay]
      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
    \end{tikzpicture}
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{General weighting strategies}

\begin{enumerate}
\item Pretest using weighted Bonferroni or conventional graphical
  approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
  the remaining hypotheses $I$
\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$ 
\item $I \gets I \setminus R$
\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
  \mathscr{H} - 1\}$ 
\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
\item Repeat until $\rho \mathscr {H} = 0$
\item Reject $R$ and $I \setminus A$
\end{enumerate}
\end{frame}
\section{Matriods from general families of hypotheses}

\begin{frame}
\frametitle{Matroid conditions}
\end{frame}

\begin{frame}
\frametitle{Flats <-> Weakliy exhaustive sets}
\end{frame}

\begin{frame}
\frametitle{Flats <-> Strictly exhaustive sets}
\end{frame}
\section{Summary \& Outlook}

\begin{frame}
\frametitle{Summary}
\end{frame}

\begin{frame}
\frametitle{Outlook}

\end{frame}

\bibliographystyle{alpha}
\bibliography{\parentd paper/cleanrefs}
\end{document}