### 2013-08-15 presentation/full4ext.tex
added a couple of graph files, set up the pictures for qva example

\begin{tikzpicture}[scale=.5]

\node (H1) at (225bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
\node (H2) at (425bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
\node (H3) at (425bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
\node (H4) at (225bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
\draw [line width=1pt] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);
\draw [line width=1pt] (H1) to (275bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);
\draw [line width=1pt] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
\draw [line width=1pt] (H3) to (326bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);
\draw [line width=1pt] (H4) to (225bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
\draw [line width=1pt] (H4) to (374bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);

\node (H1) at (525bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
\node (H2) at (725bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
\node (H3) at (725bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
\node (H4) at (525bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
\draw [line width=1pt] (H1) to (625bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);
\draw [line width=1pt] (H1) to (575bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);
\draw [line width=1pt] (H2) to (725bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
\draw [line width=1pt] (H3) to (626bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);
\draw [line width=1pt] (H4) to (525bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
\draw [line width=1pt] (H4) to (674bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);

\end{tikzpicture}