### 2013-08-15 presentation/lsd2.tex

\begin{tikzpicture}[scale=.5]
\node (Pl) at (25bp,-225bp)[draw,circle,fill=white] {$Pl$};
\node (Dl) at (175bp,-175bp)[draw,circle,fill=white] {$Dl$};
\node (Dh) at (225bp,-25bp)[draw,circle,fill=white] {$Dh$};
\draw [line width=1pt] (Pl.363) arc(273:290:209bp) node[fill=blue!20] {$H_{2}$} arc(290:302:209bp) to (Dl);
\draw [line width=1pt] (Pl.70) arc(160:145:282bp) node[fill=blue!20] {$H_{1}$} arc(145:112:282bp) to (Dh);
\draw [line width=1pt] (Dl.416) arc(326:340:212bp) node[fill=blue!20] {$H_{3}$} arc(340:355:212bp) to (Dh);
\end{tikzpicture}