### 2013-08-19 presentation/presentation.tex
incorporated willi's changes

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{ulem}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\em i.e.},~}
\newcommand{\eg}{{\em e.g.},~}
\newcommand{\parentd}{../}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage

\title{Matroids from families of null hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}
\frametitle{Matroids}

\maketitle

\begin{frame}
\frametitle{Outline}
\setcounter{tocdepth}{3}
\tableofcontents
\end{frame}

\section{Logically dependent hypotheses}

\begin{frame}
\frametitle{Pairwise comparisons, some history }

Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
\begin{itemize}
  \item Parameters: $\mu_1,\mu_2,\mu_3$ ; \\ Hypotheses: $H_1:\mu_1=\mu_2,H_2:\mu_2=\mu_3;H_3:\mu_1=\mu_3$
  \item A level-$\alpha$ test of $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 : \mu_1=\mu_2=\mu_3$. followed by level-$\alpha$ tests of $H_1,H_2,H_3$ protects the FWER at level $\alpha$.
  \begin{itemize}
    \item  essentially closed test; no need to test pairwise intersection hypotheses since $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j$,
    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise comparisons, some history}

Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
\begin{itemize}
  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6$ :
    \begin{itemize}
      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
      \item Dynamic: $t^*_j \le t_j$, hence more powerful than "static" 
           \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{tabular}{ll}
  \begin{minipage}{.3\textwidth}
    \only<2->{\input{full4}}
.
  \end{minipage} &
  \begin{minipage}{.7\textwidth}
  \begin{itemize}
\item Consider $\mu_i, i \in \{1,...,m\}$

\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
  \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Max number of true hypotheses}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
   \input{shaffer}
.
  \end{minipage} &
  \begin{minipage}{.6\textwidth}
  \begin{itemize}
  \item "Static":  $t_j$ = max number of true hypotheses if $j$ are false.
  \item "Dynamic": $t^*_j $ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
  \item Example: $m=4, n=6, j = 2$ :
    \begin{itemize}
      \item  {Static: if, e.g., $H_1$ and $H_2$ are false then at most $H_4, H_5 $ and $H_6$ can be simultaneously true, hence $t_2 = 3$}
      \item  {Dynamic: if, e.g., $H_1$ and $H_6$ are false then only $H_3$ and $H_4$, or $H_3$ and $H_5$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$.}
      \end{itemize}

\end{itemize} % ends low level
  \end{minipage}
\end{tabular}
\end{frame}
%
%
%\begin{frame}
%\frametitle{Pairwise mean comparisons (two-sided)}
%\begin{itemize}
%
%\item Consider $\mu_i, i \in \{1,...,m\}$
%
%\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$
%
%\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
%\begin{itemize}
%\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
%\item edges correspond to hypotheses, \emph{i.e.}, $E =
%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
%      j$, and $e_l \neq e_k$ for $l \neq k$
%\end{itemize}
%\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
%\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
%  \in I\}$
%\end{itemize} % ends low level
%\end{frame}
%
%\begin{frame}
%\frametitle{Example 1: 4 Groups}
%\input{full4ext}
%\end{frame}
%

% \begin{frame}
% \frametitle{Example 1: 4 Means}
% \input{full4}
% \end{frame}

%
%\begin{frame}
%\frametitle{General contrast tests (two-sided)}
%\begin{itemize}
%
%\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$
%
%\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
%  c_{ij} = 0$
%\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
%  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
%    I}$
%\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
%\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$
%
%\item Elementary hypotheses are represented by column vectors,
%  intersection hypotheses by sub-matrices of the contrast matrix
%
%\end{itemize} % ends low level
%\end{frame}
%

%
%\begin{frame}
%\frametitle{Example 2: Contrast test}
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.45\textwidth}
%    All pairwise comparisons of 4 means
%    \begin{displaymath}
%    \bs{C}^t = \left(
%      \begin{array}{llll}
%      -1 & 1 & 0 & 0 \\
%      -1 & 0 & 1 & 0 \\
%      -1 & 0 & 0 & 1 \\
%      0 & -1 & 1 & 0 \\
%      0 & -1 & 0 & 1 \\
%      0 & 0 & -1 & 1
%    \end{array}
%    \right)
%    \end{displaymath}
%  \end{minipage} &
%  \begin{minipage}{.55\textwidth}
%    Modified Williams Contrasts for Dose Response
%    \begin{displaymath}
%      \bs{C}^t=\left(
%      \begin{array}{llll}
%        1 & 0 & 0 & -1\\
%        1 & 0 &-.51 & -.49 \\
%        1 & -.35 & -.33 & -.32\\
%        .51 & .49 & -.51 & -.48\\
%        .51 & .49 & 0 &  -1\\
%        .35 & .33 & .32& -1
%      \end{array}\right)
%    \end{displaymath}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Dependent null hypotheses}
%
%\begin{itemize}
%\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
%  dependence relation $\sim_H$ on $\mathcal{P}(S)$
%\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
%\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
%  H_{I\setminus i}$ for all $i \in I$
%\pause
%\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
%\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
%  column rank
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\section{Matroids}
%
%
%
%\begin{frame}
%\frametitle{Matroids}
%\begin{itemize}
%
%\item Matroids generalize the structures of graphs and vector spaces,
%  \eg cycles, bases, independent sets
%\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
%  structure
%\item Cryptomorphism: many ways to define a matroid:
%  \begin{itemize}
%  \item From a graph perspective: Circuit axioms.
%  \item From a vector space perspective: Independence axioms.
%  \end{itemize}
%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Circuit axioms}
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \only<2->{\input{cycle}}
%.
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
%    \begin{enumerate}
%    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
%    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
%      $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
%    \end{enumerate}
%    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%% \item \textbf{/Some illustration hereabouts/}
%
%% \end{itemize} % ends low level
%% \end{frame}
%
%
%\begin{frame}
%\frametitle{Independence axioms}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%  \only<1>{\input{independent}}
%  \only<2>{\input{cycle}}
%  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
%  holds
%  \begin{enumerate}
%  \item $\emptyset \in \mathscr{I}$,
%  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
%    \mathscr{I}$,
%  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
%    \setminus J$ such that $J \cup i \in \mathscr{I}$,
%  \end{enumerate}
%  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
%  on $S$. This defines an independence relation $i \nsim_M I$ iff
%  $I\cap i \in \mathscr{I}$.
%
%  \end{minipage}
%\end{tabular}
%
%
%
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Dependence axioms}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \input{dependent}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%  Relation $\sim_M$ on elements and subsets of $S$,
%  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
%  \begin{enumerate}
%  \item $y \sim Y$ for all $y \in Y$
%  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
%    then $y \sim x \cup (Y \setminus y)$
%  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
%  \end{enumerate}
%  then $\sim_M$ defines a matroid on $S$.
%  \end{minipage}
%\end{tabular}
%
%
%
%\end{frame}
%
%\section{Matroids from logically dependent hypotheses}
%\begin{frame}
%  \frametitle{Dependent hypotheses}
%  \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
%    \begin{itemize}
%    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
%    \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
%      $\sim_H = \sim_M$?
%    \end{itemize}
%  \end{block}
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Matroids from logically dependent hypotheses}
%
%  \begin{block}{Two-sided pariwise comparisons}
%    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
%    S$, an elementary hypotheses $H_j$ depends on intersection
%    hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
%    the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by
%    $\mathscr{G}_I$.
%  \end{block}
%
%  \begin{block}{Two-sided linear contrast tests}
%    For a family of two-sided general contrast tests $H_i$, $i \in S$,
%    an elementary hypothesis $H_j$ depends on intersection hypotheses
%    $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
%    dependent on the column vectors of the matrix $\bs{C}_I$ in the
%    vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
%  \end{block}
%
%\end{frame}
%
%
%\begin{frame}
%  \frametitle{Matroids from null hypotheses}
%  {\Large What is it good for?\\
%  \invisible<1>{Structure}}
%\end{frame}
%
%\begin{frame}
%\frametitle{Maximally independent sets}
%  \begin{itemize}
%  \item Bases are maximally independent sets of $S$
%  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
%  \item Circuits are minimally dependent sets, \ie each subset of $C$
%    with rank $\rho C - 1$ is a maximally independent set of $C$
%  \end{itemize}
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Maximally dependent sets}
%\begin{itemize}
%\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
%\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
%\item Hyperplanes are flats of rank $\rho S - 1$
%\item All flats are given by closing hyperplanes under intersection
%\item Circuits in $M^*$ are the complements of hyperplanes in $M$
%\end{itemize} % ends low level
%\end{frame}
%
%
%
%\begin{frame}{Examples}
%\framesubtitle<1>{Independent set}
%\framesubtitle<2>{Base}
%\framesubtitle<4>{Dependent set}
%\framesubtitle<3>{Cycle}
%\framesubtitle<5>{Closure}
%\only<1>{\input{independent}}
%\only<2>{\input{base}}
%\only<4>{\input{dependent}}
%\only<3>{\input{cycle}}
%\only<5>{\input{closure}}
%\end{frame}
%
%
%
%
%
%
%
%
%\begin{frame}
%\frametitle{Flats}
%\begin{itemize}
%
%\item The flats of a matroid are closed under intersection
%
%\item Every elementary hypothesis is a flat
%\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
%\item Define a stepwise procedure on the flats of $M$
%
%
%\end{itemize} % ends low level
%\end{frame}
%
%\begin{frame}
%  \frametitle{Stepwise test procedure}
%  \begin{block}{Multiple testing procedure for logically related hypotheses}
%    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
%    defines a matroid on $M$ on $S$. Then a test procedure that
%    rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
%    \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
%    controls the FWER in the strong sense.
%  \end{block}
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Stepwise multiple testing procedure on the lattice of
%    flats}
%%
%\begin{tabular}{ll}
%  \begin{minipage}{.3\textwidth}
%    \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
%    \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
%    \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
%    \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
%    {\tiny wikipedia:Partition\_of\_a\_Set}
%  \end{minipage} &
%  \begin{minipage}{.7\textwidth}
%  \begin{itemize}
%  \item All flats are given by intersections of hyperplanes
%    $\rightarrow$ knowing all hyperplanes remaining flats are given by
%    intersection of index sets
%  \item Hyperplanes can by found by algorithms that enumerate the
%    circuits of the dual matroid \cite{boros2003algorithms}
%  \item The partial order of flats defined by inclusion defines a
%    geometric lattice
%  \item<2-> Testing only a subset of elementary hypotheses corresponds
%    to intervals
%  \item<3-> Rejection of elementary hypotheses removes intervals
%  \end{itemize}
%
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%\begin{frame}{Matroid procedures for pairwise mean comparisons}
%\begin{tabular}{ll}
%  \begin{minipage}{.3\textwidth}
%    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
%     {\tiny wikipedia:Bell\_number}
%  \end{minipage} &
%  \begin{minipage}{.7\textwidth}
%  \begin{itemize}
%  \item Two-sided comparisons of $m$ means
%  \item The cycle matroid of the graph representation
%    provides equivalent dependence structure
%  \item The flats of a (simple) complete graph are
%    isomporphic to the partitions of a set of size $m$
%  \item Efficient algorithms to generate all exhaustive sets by
%    enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new}
%  \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
%    51, 202, 876, 4139, ...
%  \end{itemize}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%
%
%
%\begin{frame}
%\frametitle{Connections to existing procedures}
%
%\textbf{Old wine in new bottles ((c) L. Hothorn)}
%\begin{itemize}
%\item \cite{shaffer1986modified} represents pairwise comparisons
%  by set partitions
%\item \cite{bergmann1988improvements} define the concepts of an
%  exhaustive sets
%\item \cite{bernhard1991computergestuetzte} shows properties of
%  exhaustive sets that are basically {\em closure axioms} of a matroid
%  and provides algorithms to check each intersection hypotheses if is
%  a flat using rank function.
%\item \cite{westfall1997multiple} generates flats using dependence
%  arguments in vector spaces.
%\item \cite{weichert2000robuste} suggest an MTP based on clique
%  decomposition of the graph, cliques in the graph are flats
%\item \cite{westfall2007multiple} provide a branch and bound algorithm
%  to generate all flats. They note that for pairwise comparisons all
%  exhaustive hypotheses are {\em partition hypotheses}
%\end{itemize}
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Matroid representation}
%  \begin{itemize}
%  \item proofs are shorter (more elegant),
%  \item efficient algorithms exist to find the flats of matroids,
%  \item {\em Scum principle}\cite{welsh2010matroid} everything
%    important happens at the top (hyperplanes)
%  \item we may define procedures using independent sets,
%  \item \ldots{}
%  \end{itemize}
%\end{frame}
%
%\section{Graphical approaches for logically dependent hypotheses}
%
%
%
%\begin{frame}
%\frametitle{Reminder: Graphical approaches}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \begin{tikzpicture}[remember picture,overlay]
%      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
%    \end{tikzpicture}
%
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    \begin{itemize}
%    \item Provides weights $w_{i,I}$ for each index $i$ and each
%      non-empty subset $I \subset S$
%    \item Weights are used to define a closed test procedure of
%      weighted intersection hypotheses tests
%    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
%    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
%    \end{itemize}
%  \end{minipage}
%\end{tabular}
%\end{frame}
%
%
%\begin{frame}
%  \frametitle{Example: Hierarchical LSD}
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \input{lsd2.tex}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    \begin{itemize}
%    \item 3 Hypotheses:
%      \begin{enumerate}
%      \item Placebo vs. High Dose
%      \item Placebo vs. Low Dose
%      \item High Dose vs. Low Dose
%      \end{enumerate}
%    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
%    \item Test $H_1$ at level $\alpha$
%    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
%      simultaneously
%    \item Can we generalize this?
%    \end{itemize}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Fixed sequence test of pairwise mean comparisons}
%\begin{itemize}
%\item {\bf Picture of hierarchical test hereabouts}
%\end{itemize}
%
%\begin{block}{Theorem}
%  Assume that $M(\mathscr{H})$ neither contains loops nor
%  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
%  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
%  to be tested at level $\alpha_r$ according to a prefixed order.  If
%  for some $1 \leq s \leq t$ holds
%  \begin{equation}
%    \label{eq:cond.hierarchical}
%    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
%  \end{equation}
%  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
%  at full level $\alpha_ra$.
%\end{block}
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Example: Multiple doses}
%
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \begin{tikzpicture}[overlay]
%      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
%    \end{tikzpicture}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    \begin{tikzpicture}[remember picture,overlay]
%      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
%    \end{tikzpicture}
%  \end{minipage}
%\end{tabular}
%\end{frame}
%
%\begin{frame}
%  \frametitle{General weighting strategies}
%  \framesubtitle{QVA Example}
%
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%    \begin{tikzpicture}[remember picture,overlay]
%      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
%    \end{tikzpicture}
%
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    \begin{tikzpicture}[remember picture,overlay]
%      \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
%  \end{tikzpicture}
%  \end{minipage}
%\end{tabular}
%\end{frame}
%
%\begin{frame}
%\frametitle{General weighting strategies}
%
%\begin{enumerate}
%\item Pretest using weighted Bonferroni or conventional graphical
%  approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
%\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
%  the remaining hypotheses $I$
%\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
%\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$
%\item $I \gets I \setminus R$
%\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
%  \mathscr{H} - 1\}$
%\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
%\item Repeat until $\rho \mathscr {H} = 0$
%\item Reject $R$ and $I \setminus A$
%\end{enumerate}
%\end{frame}
%\section{Matriods from general families of hypotheses}
%
%
%
%\begin{frame}
%\frametitle{General Results}
%\begin{itemize}
%\item We give sufficient (and probably necessary) conditions for a
%  general family of hypotheses $\mathscr{H}$ to define a matroid
%\item We give sufficient (and probably necessary) conditions on
%  $\mathscr{H}$ for $\mathscr{L}$ to be the set of strictly exhaustive
%  sets
%\item Systems with one sided hypotheses, and parallel hypotheses (\eg
%  combined non-inferiority/superiority) do not give matroids with
%  $\mathscr{L}$ all strictly exhaustive
%\end{itemize}
%
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{One sided hypotheses}
%  \begin{block}{One-sided linear contrast hypotheses}
%    A family of {\em one sided linear contrast hypotheses} is defined
%    by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and
%    vector $\bs{b} \in \mathbb{R}^n$ that define a  family of null
%    hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu}
%    \leq b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th
%    column of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
%  \end{block}
%\end{frame}
%
%\begin{frame}
%  \frametitle{Hyperplane arrangement}
%  \begin{itemize}
%  \item Each hypothesis $H_i$ corresponds to a hyperplane $h_i =
%    \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$
%  \item Each hyperplane $h_i$ divides $\mathbb{R}^m$ into two regions
%    \begin{itemize}
%    \item $h_i^+ = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} <
%      b_i\}$
%    \item $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x}
%      > b_i\}$
%    \end{itemize}
%  \item Defines a sign pattern
%    \begin{equation}
%      \label{eq:sign}
%      \delta_i(\bs{x}) = \left\{
%        \begin{array}{ll}
%          + & \text{if } \bs{x} \in h_i^+\\
%          0 & \text{if } \bs{x} \in h_i \\
%          - & \text{if } \bs{x} \in h_i^-\\
%        \end{array}\right.
%    \end{equation}
%  \item $\delta_i$ is $-$ if $H_i$ false at $x$, or $0$ or $+$
%    otherwise. $(\delta_1,...,\delta_n)$ tells us to which combination
%    of true and false null hypotheses some $\bs{x}$ belongs.
%
%  \item Set of all sign patterns defined by $\mathscr{H}$:
%    \begin{equation}
%      \label{eq:faces}
%      \mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
%    \end{equation}
%  \end{itemize}
%\end{frame}
%
%\begin{frame}
%  \frametitle{Example}
%
%\begin{tabular}{ll}
%  \begin{minipage}{.4\textwidth}
%  \includegraphics[height=1\textwidth]{spheresfukuda.eps}
%  \end{minipage} &
%  \begin{minipage}{.6\textwidth}
%    \begin{itemize}
%    \item 4 contrast hypotheses about 3 means
%    \item Cut the hyperplanes with the sphere in $\R^3$ (makes
%      visualisation easier)
%    \item Cells of the sphere arrangment correspond to sign patterns
%      of $\mathscr{H}$
%    \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
%    \end{itemize}
%  \end{minipage}
%\end{tabular}
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Stepwise multiple testing procedure for one-sided linear
%    contrast hypotheses}
%
%  Consider a family of general contrast hypotheses $\mathscr{H} =
%  \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
%  $\mathscr{F}(\mathscr{H}) = \{ \bs{\delta}(\bs{x}): \bs{x} \in
%  \R^m\}$ denote the corresponding set of sign patterns. Consider a
%  set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
%  which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
%  corresponding decision rule for each elementary hypotheses $H_i$ is
%  given by
%  \begin{equation}
%    \label{eq:signtest}
%    \psi_i = \min_{F: F_i \geq 0} \varphi_F,
%  \end{equation}
%  which controls the FWER at level $\alpha$ in the strong sense.
%
%
%\end{frame}
%
%
%\begin{frame}
%  \frametitle{Computation of $\mathscr{F}$}
%  The question remains how to compute $\mathcal{F}$. For this observe
%that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
%just the face lattice of the corresponding oriented
%matroid. Consequently, $\mathcal{F}$ can be found using algorithms
%for arrangement construction from polyhedral programming
%\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
%polynomial in time and input size and compact in space!
%
%\end{frame}
%
%\begin{frame}
%  \frametitle{Example}
%  3 ordered means
%\end{frame}
%
%\section{Summary \& Outlook}
%
%
%
%\begin{frame}
%\frametitle{Summary}
%\begin{itemize}
%\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
%\end{itemize}
%\end{frame}
%
%
%
%\begin{frame}
%\frametitle{Outlook}
%\begin{itemize}
%\item MCPmod on \sout{steroids} matroids
%\end{itemize}
%\end{frame}
%
%\bibliographystyle{alpha}
%\bibliography{\parentd paper/cleanrefs}
\end{document}