### 2013-08-19 presentation/presentation.tex
cleaned up the first couple of slides

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{ulem}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\sl i.e.},~}
\newcommand{\eg}{{\sl e.g.},~}
\newcommand{\parentd}{../}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage
\setbeamertemplate{blocks}[rounded][shadow=true]

\title{Matroids \& testing in families of dependent hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}
\frametitle{Matroids}

\maketitle

\begin{frame}
\frametitle{Outline}
\setcounter{tocdepth}{3}
\tableofcontents
\end{frame}

\section{Algebraically dependent hypotheses}

\begin{frame}
\frametitle{Some history:}
\framesubtitle{Testing hypotheses of no differences in pairwise
  comparisons of 3 treatments}

% Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):

\begin{block}{$H_1:\mu_1=\mu_2,\; H_2:\mu_2=\mu_3,\;H_3:\mu_1=\mu_3$}
  \begin{enumerate}
  \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
    \mu_1=\mu_2=\mu_3$ at level $\alpha$,
  \item test each $H_1,H_2,H_3$ at level $\alpha$,
  \end{enumerate}
  protects the FWER at level $\alpha$.
\end{block}
\pause
  \begin{itemize}
    \item  essentially closed test, since %; no need to test pairwise
                                %intersection hypotheses since 
        $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j,$

    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}

\end{frame}

\begin{frame}
\frametitle{Some history}
\framesubtitle{Testing hypotheses of no difference in pairwise
    comparisons of $m>3$ treatments:}

  Modified Sequentially Rejective
  Multiple test procedures \cite{shaffer1986modified} % Example: Modified Bonferroni-Holm
      \begin{itemize}
      \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
      \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test
        remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ =
        max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are
        false.
      \end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{tabular}{ll}

  \begin{minipage}{.5\textwidth}
    \input{full4}
  \end{minipage} 
  \begin{minipage}{.5\textwidth}
    \begin{block}{Graphical representation}
      \begin{itemize}
      \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq
        m*(m-1)/2$

      \item Undirected graph %$\mathscr{G} = (V,E)$:
        \begin{itemize}
        \item vertices correspond to means, 
          %\emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
        \item edges correspond to hypotheses, 
          % \emph{i.e.}, $E = \left\{e_1,...,e_n\} \right\}$, $e_\ell = \{\mu_i,\mu_j\}$, $i
          % \neq j$, and $e_\ell \neq e_k$ for $\ell \neq k$
        \end{itemize}
      \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
      \item $I$ induces a subgraph %$\mathscr{G}_I$ 
        on the edge set
        %$E_I=\{e_i: i \in I\}$
      \end{itemize} % ends low level
    \end{block}
  \end{minipage}

\end{tabular}
\end{frame}

% \begin{block}{Example: $m=4, n=6$ :}
%       \begin{itemize}
%       \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
%       \item Dynamic: $t^*_j \le t_j$, hence more powerful than
%         "static"
%       \end{itemize}
%   \end{block}

\begin{frame}
\frametitle{Example: Modified Bonferroni Holm \cite{shaffer1986modified}}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}

    \input{shaffer}
\end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{block}{$m=4, n=6, \only<2->{j=2}$}
      % \begin{itemize}
      % \item "Static": $t_j$ = max number of true hypotheses if $j$ are
      %   false.
      % \item "Dynamic": $t^*_j $ = max. number of true hypotheses if
      %   $H_{i_1},...,H_{i_j}$ are false.
%      \item Example:  \only<3->{$ j = 2$ :}
        \begin{description}
        \item<3->[Static:]  $H_1$ and $H_2$ are false
            $\Rightarrow$ $t_2 = 3$ (at most $H_4, H_5 $ and $H_6$ can be
            simultaneously true)
        \item<7->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
          $t^*_2 = 2 < t_2 =3$ (only $H_3$, $H_4$, or $H_2$, $H_5$ can be
            simultaneously true)
        \end{description}

 %     \end{itemize} % ends low level
    \end{block}
  \end{minipage}
\end{tabular}
\end{frame}

% \begin{frame}
% \frametitle{Pairwise mean comparisons (two-sided)}
% \begin{itemize}

% \item Consider $\mu_i, i \in \{1,...,m\}$

% \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

% \item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
% \begin{itemize}
% \item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
% \item edges correspond to hypotheses, \emph{i.e.}, $E =
%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
%      j$, and $e_l \neq e_k$ for $l \neq k$
% \end{itemize}
% \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
% \item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
%  \in I\}$
% \end{itemize} % ends low level
% \end{frame}

% \begin{frame}
% \frametitle{Example 1: 4 Groups}
% \input{full4ext}
% \end{frame}

% \begin{frame}
% \frametitle{Example 1: 4 Means}
% \input{full4}
% \end{frame}

\begin{frame}
\frametitle{General contrast tests (two-sided)}
\begin{itemize}

\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$

\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
 c_{ij} = 0$
\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
 with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
   I}$
\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$

\item Elementary hypotheses are represented by column vectors,
 intersection hypotheses by sub-matrices of the contrast matrix

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 2: Contrast test}

\begin{tabular}{ll}
 \begin{minipage}{.44\textwidth}
   \begin{block}{\footnotesize Pairwise comparisons of 4 means}
     {\footnotesize
       \begin{displaymath}
         \bs{C}^t = \left(
           \begin{array}{rrrr}
             -1 & 1 & 0 & 0 \\
             -1 & 0 & 1 & 0 \\
             -1 & 0 & 0 & 1 \\
             0 & -1 & 1 & 0 \\
             0 & -1 & 0 & 1 \\
             0 & 0 & -1 & 1
           \end{array}
         \right)
       \end{displaymath}}
   \end{block}
 \end{minipage} &
 \begin{minipage}{.56\textwidth}
\pause
   \begin{block}{\footnotesize Modified Williams Contrasts for Dose Response \cite{bretz2010multiple}}
      {\footnotesize
       \begin{displaymath}
         \bs{C}^t=\left(
           \begin{array}{rrrr}
             1 & 0 & 0 & -1\\
             1 & 0 &-.51 & -.49 \\
             1 & -.35 & -.33 & -.32\\
             .51 & .49 & -.51 & -.48\\
             .51 & .49 & 0 &  -1\\
             .35 & .33 & .32& -1
           \end{array}\right)
       \end{displaymath}}
   \end{block}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependent null hypotheses}

\begin{itemize}
\item $H_i$ depends on $H_I$: $H_{I \cup i} = H_{I}$
\item induces a dependence relation $i \sim_H I$ on index sets $I \in \mathcal{P}(S)$
\item $I$ independent iff for all $i
  \in I$ $i \nsim_H I\setminus i$ (\ie $H_{I} \subsetneq H_{I\setminus i}$)
\pause
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
 column rank
\end{itemize} % ends low level
\end{frame}

\section{Matroids}

\begin{frame}
\frametitle{Matroids}
\begin{itemize}

\item Matroids generalize the structures of graphs and vector spaces,
 \eg cycles, bases, independent sets
\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
 structure
\item {\em Cryptomorphism} \ie many ways to define a matroid:
 \begin{itemize}
 \item From a graph perspective: Circuit axioms.
 \item From a vector space perspective: Independence axioms.
 \end{itemize}
%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Circuit axioms}

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Circuits of a matroid $M$ on $S$:}
       A collection $\mathscr{C}$ of subsets of $S$
       for which holds
     \begin{enumerate}
     \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
     \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$
       then $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
     \end{enumerate}
   \end{block}

 \end{minipage} &

 \begin{minipage}{.5\textwidth}
   \only<2->{\input{cycle}}
 \end{minipage} 
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Independence axioms}

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Independent sets of a matroid $M$ on $S$}
     A a collection $\mathscr{I}$ of subsets of $S$ for which
     holds
     \begin{enumerate}
     \item $\emptyset \in \mathscr{I}$,
     \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
       \mathscr{I}$,
     \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in
       I \setminus J$ such that $J \cup i \in \mathscr{I}$,
     \end{enumerate}
   \end{block}
 \end{minipage} &

 \begin{minipage}{.5\textwidth}
 \only<1>{\input{independent}}
 \only<2>{\input{cycle}}
 \invisible<1>{\alert{\footnotesize minimally dependent set!}}
 \end{minipage} 
\end{tabular}

\end{frame}

\begin{frame}
  \frametitle{}
  \begin{itemize}
  \item Circuits are minimally dependent sets (remove any element to
    get an independent set)
  \item Independent sets are sets for which no subset of elements is a
    circuit
  \item Maximally independent sets are bases (Base axioms)
  \item Cardinality of largest independent subset is the rank of a set
    (Rank axioms)
  \item Maximally dependent sets are called flats (Closure axioms)
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Dependence axioms}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \input{dependent}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
 Relation $\sim_M$ on elements and subsets of $S$,
 if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
 \begin{enumerate}
 \item $y \sim Y$ for all $y \in Y$
 \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
   then $y \sim x \cup (Y \setminus y)$
 \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
 \end{enumerate}
 then $\sim_M$ defines a matroid on $S$.
 \end{minipage}
\end{tabular}

\end{frame}

\section{Matroids from algebraically dependent hypotheses}
\begin{frame}
 \frametitle{Dependent hypotheses}
 \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
   \begin{itemize}
   \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
   \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
     $\sim_H = \sim_M$?
   \end{itemize}
 \end{block}
\end{frame}

\begin{frame}
\frametitle{Matroids from algebraically dependent hypotheses}

 \begin{block}{Two-sided pariwise comparisons}
   For a family of two-sided pairwise mean comparisons $H_i$, $i \in
   S$, an elementary hypotheses $H_j$ depends on intersection
   hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
   the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by
   $\mathscr{G}_I$.
 \end{block}

 \begin{block}{Two-sided linear contrast tests}
   For a family of two-sided general contrast tests $H_i$, $i \in S$,
   an elementary hypothesis $H_j$ depends on intersection hypotheses
   $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
   dependent on the column vectors of the matrix $\bs{C}_I$ in the
   vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
 \end{block}

\end{frame}

\begin{frame}
 \frametitle{Matroids from null hypotheses}
 {\Large What is it good for?\\
 \invisible<1>{Structure}}
\end{frame}

\begin{frame}
\frametitle{Maximally independent sets}
 \begin{itemize}
 \item Bases are maximally independent sets of $S$
 \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
 \item Circuits are minimally dependent sets, \ie each subset of $C$
   with rank $\rho C - 1$ is a maximally independent set of $C$
 \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Maximally dependent sets}
\begin{itemize}
\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
\item Hyperplanes are flats of rank $\rho S - 1$
\item All flats are given by closing hyperplanes under intersection
\item Circuits in $M^*$ are the complements of hyperplanes in $M$
\end{itemize} % ends low level
\end{frame}

\begin{frame}{Examples}
\framesubtitle<1>{Independent set}
\framesubtitle<2>{Base}
\framesubtitle<4>{Dependent set}
\framesubtitle<3>{Cycle}
\framesubtitle<5>{Closure}
\only<1>{\input{independent}}
\only<2>{\input{base}}
\only<4>{\input{dependent}}
\only<3>{\input{cycle}}
\only<5>{\input{closure}}
\end{frame}

\begin{frame}
\frametitle{Flats}
\begin{itemize}

\item The flats of a matroid are closed under intersection

\item Every elementary hypothesis is a flat
\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
\item Define a stepwise procedure on the flats of $M$

\end{itemize} % ends low level
\end{frame}

\begin{frame}
 \frametitle{Stepwise test procedure}
 \begin{block}{Multiple testing procedure for algebraically related hypotheses}
   Assume that $\mathscr{H}$ with dependence relation $\sim_H$
   defines a matroid on $M$ on $S$. Then a test procedure that
   rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
   \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
   controls the FWER in the strong sense.
 \end{block}

\end{frame}

\begin{frame}
 \frametitle{Stepwise multiple testing procedure on the lattice of
   flats}

\begin{tabular}{ll}
 \begin{minipage}{.3\textwidth}
   \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
   \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
   \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
   \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
   {\tiny wikipedia:Partition\_of\_a\_Set}
 \end{minipage} &
 \begin{minipage}{.7\textwidth}
 \begin{itemize}
 \item All flats are given by intersections of hyperplanes
   $\rightarrow$ knowing all hyperplanes remaining flats are given by
   intersection of index sets
 \item Hyperplanes can by found by algorithms that enumerate the
   circuits of the dual matroid \cite{boros2003algorithms}
 \item The partial order of flats defined by inclusion defines a
   geometric lattice
 \item<2-> Testing only a subset of elementary hypotheses corresponds
   to intervals
 \item<3-> Rejection of elementary hypotheses removes intervals
 \end{itemize}

 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}{Matroid procedures for pairwise mean comparisons}
\begin{tabular}{ll}
 \begin{minipage}{.3\textwidth}
   \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
    {\tiny wikipedia:Bell\_number}
 \end{minipage} &
 \begin{minipage}{.7\textwidth}
 \begin{itemize}
 \item Two-sided comparisons of $m$ means
 \item The cycle matroid of the graph representation
   provides equivalent dependence structure
 \item The flats of a (simple) complete graph are
   isomporphic to the partitions of a set of size $m$
 \item Efficient algorithms to generate all exhaustive sets by
   enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new}
 \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
   51, 202, 876, 4139, ...
 \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Connections to existing procedures}

\textbf{Old wine in new bottles ((c) L. Hothorn)}
\begin{itemize}
\item \cite{shaffer1986modified} represents pairwise comparisons
 by set partitions
\item \cite{bergmann1988improvements} define the concepts of an
 exhaustive sets
\item \cite{bernhard1991computergestuetzte} shows properties of
 exhaustive sets that are basically {\em closure axioms} of a matroid
 and provides algorithms to check each intersection hypotheses if is
 a flat using rank function.
\item \cite{westfall1997multiple} generates flats using dependence
 arguments in vector spaces.
\item \cite{weichert2000robuste} suggest an MTP based on clique
 decomposition of the graph, cliques in the graph are flats
\item \cite{westfall2007multiple} provide a branch and bound algorithm
 to generate all flats. They note that for pairwise comparisons all
 exhaustive hypotheses are {\em partition hypotheses}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Matroid representation}
 \begin{itemize}
 \item proofs are shorter (more elegant),
 \item efficient algorithms exist to find the flats of matroids,
 \item {\em Scum principle}\cite{welsh2010matroid} everything
   important happens at the top (hyperplanes)
 \item we may define procedures using independent sets,
 \item \ldots{}
 \end{itemize}
\end{frame}

\section{Graphical approaches for algebraically dependent and
  contextually related hypotheses}

\begin{frame}
\frametitle{Reminder: Graphical approaches}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
   \end{tikzpicture}

 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item Provides weights $w_{i,I}$ for each index $i$ and each
     non-empty subset $I \subset S$
   \item Weights are used to define a closed test procedure of
     weighted intersection hypotheses tests
   \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
   \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
   \end{itemize}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
 \frametitle{Example: Hierarchical LSD}
\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \input{lsd2.tex}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item 3 Hypotheses:
     \begin{enumerate}
     \item Placebo vs. High Dose
     \item Placebo vs. Low Dose
     \item High Dose vs. Low Dose
     \end{enumerate}
   \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
   \item Test $H_1$ at level $\alpha$
   \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
     simultaneously
   \item Can we generalize this?
   \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
 \frametitle{Fixed sequence test of pairwise mean comparisons}
\begin{itemize}
\item {\bf Picture of hierarchical test hereabouts}
\end{itemize}

\begin{block}{Theorem}
 Assume that $M(\mathscr{H})$ neither contains loops nor
 parallels. Assume that hypotheses $H_1$ through $H_r$ have been
 rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
 to be tested at level $\alpha_r$ according to a prefixed order.  If
 for some $1 \leq s \leq t$ holds
 \begin{equation}
   \label{eq:cond.hierarchical}
   \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
 \end{equation}
 hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
 at full level $\alpha_ra$.
\end{block}

\end{frame}

\begin{frame}
 \frametitle{Example: Multiple doses}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[overlay]
     \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
   \end{tikzpicture}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
   \end{tikzpicture}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
 \frametitle{General weighting strategies}
 \framesubtitle{QVA Example}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
   \end{tikzpicture}

 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
 \end{tikzpicture}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{General weighting strategies}

\begin{enumerate}
\item Pretest using weighted Bonferroni or conventional graphical
 approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
 the remaining hypotheses $I$
\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$
\item $I \gets I \setminus R$
\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
 \mathscr{H} - 1\}$
\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
\item Repeat until $\rho \mathscr {H} = 0$
\item Reject $R$ and $I \setminus A$
\end{enumerate}
\end{frame}
\section{Matriods from general families of hypotheses}

\begin{frame}
\frametitle{General Results}
\begin{itemize}
\item We give sufficient (and probably necessary) conditions for a
 general family of hypotheses $\mathscr{H}$ to define a matroid
\item We give sufficient (and probably necessary) conditions on
 $\mathscr{H}$ for $\mathscr{L}$ to be the set of strictly exhaustive
 sets
\item Systems with one sided hypotheses, and parallel hypotheses (\eg
 combined non-inferiority/superiority) do not give matroids with
 $\mathscr{L}$ all strictly exhaustive
\end{itemize}

\end{frame}

\begin{frame}
 \frametitle{One sided hypotheses}
 \begin{block}{One-sided linear contrast hypotheses}
   A family of {\em one sided linear contrast hypotheses} is defined
   by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and
   vector $\bs{b} \in \mathbb{R}^n$ that define a  family of null
   hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu}
   \leq b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th
   column of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
 \end{block}
\end{frame}

\begin{frame}
 \frametitle{Hyperplane arrangement}
 \begin{itemize}
 \item Each hypothesis $H_i$ corresponds to a hyperplane $h_i =
   \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$
 \item Each hyperplane $h_i$ divides $\mathbb{R}^m$ into two regions
   \begin{itemize}
   \item $h_i^+ = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} <
     b_i\}$
   \item $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x}
     > b_i\}$
   \end{itemize}
 \item Defines a sign pattern
   \begin{equation}
     \label{eq:sign}
     \delta_i(\bs{x}) = \left\{
       \begin{array}{ll}
         + & \text{if } \bs{x} \in h_i^+\\
         0 & \text{if } \bs{x} \in h_i \\
         - & \text{if } \bs{x} \in h_i^-\\
       \end{array}\right.
   \end{equation}
 \item $\delta_i$ is $-$ if $H_i$ false at $x$, or $0$ or $+$
   otherwise. $(\delta_1,...,\delta_n)$ tells us to which combination
   of true and false null hypotheses some $\bs{x}$ belongs.

 \item Set of all sign patterns defined by $\mathscr{H}$:
   \begin{equation}
     \label{eq:faces}
     \mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
   \end{equation}
 \end{itemize}
\end{frame}

\begin{frame}
 \frametitle{Example}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
 \includegraphics[height=1\textwidth]{spheresfukuda.pdf}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item 4 contrast hypotheses about 3 means
   \item Cut the hyperplanes with the sphere in $\R^3$ (makes
     visualisation easier)
   \item Cells of the sphere arrangment correspond to sign patterns
     of $\mathscr{H}$
   \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
   \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
 \frametitle{Stepwise multiple testing procedure for one-sided linear
   contrast hypotheses}

 Consider a family of general contrast hypotheses $\mathscr{H} =
 \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
 $\mathscr{F}(\mathscr{H}) = \{ \bs{\delta}(\bs{x}): \bs{x} \in
 \R^m\}$ denote the corresponding set of sign patterns. Consider a
 set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
 which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
 corresponding decision rule for each elementary hypotheses $H_i$ is
 given by
 \begin{equation}
   \label{eq:signtest}
   \psi_i = \min_{F: F_i \geq 0} \varphi_F,
 \end{equation}
 which controls the FWER at level $\alpha$ in the strong sense.

\end{frame}

\begin{frame}
 \frametitle{Computation of $\mathscr{F}$}
 The question remains how to compute $\mathcal{F}$. For this observe
that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
just the face lattice of the corresponding oriented
matroid. Consequently, $\mathcal{F}$ can be found using algorithms
for arrangement construction from polyhedral programming
\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
polynomial in time and input size and compact in space!

\end{frame}

\begin{frame}
 \frametitle{Example}
 3 ordered means
\end{frame}

\section{Summary \& Outlook}

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
\item Efficient algorithms from the realms of combinatorics and algebraic geometry
\item {\em Scum principle} everything important happens at the top
\item Stepwise procedure for algebraically dependent and contextually
  related hypotheses (graphical approach)
\item Promising results also for one-sided hypotheses
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outlook}
\begin{itemize}
\item Inclusion of procedures for two-sided tests in \texttt{gMCP} 
\item Implementation of one-sided problems using polyhedral
  programming libraries
\item Potential applications:
  \begin{itemize}
  \item \texttt{MCPmod} on \sout{steroids} matroids
  \item Model selection
  \item Logic regression
  \end{itemize}
\item Tests for intersection hypotheses based on maximally independent
  sets 
\end{itemize}
\end{frame}

\bibliographystyle{alpha}
\bibliography{\parentd paper/cleanrefs}
\end{document}