### 2013-08-20 presentation/shaffer.tex
added animated shaffer procedure

\begin{tikzpicture}[scale=.4]
\node (H1) at (025bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
\node (H2) at (225bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
\node (H3) at (225bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
\node (H4) at (025bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
\only<-2>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
\only<-2>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<3>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
\only<3>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=red!20] {$H_{2}$} to (H3);}
\only<-5>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<-4>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<-4>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<-4>{\draw [line width=1pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
\only<4,5>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
\only<4,5>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
\only<4,5>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=green!20] {$H_{6}$} to (H4);}

\only<6>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
\only<7>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
\only<6>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
\only<7>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);}

\only<6,7,8,9>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<6,7>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<6,7>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<6,7,8,9>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<8>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=green!20] {$H_{3}$} to (H1);}
\only<8>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
\only<9>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<9>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}

\only<10>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<10,11,12>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<10,11,12>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<10>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<11,12>{\draw [line width=2pt] (H1) to (075bp, -175bp) node[fill=green!20] {$H_{2}$} to (H3);}
\only<11,12>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}

\end{tikzpicture}