### 2013-08-20 presentation/presentation.tex
further clean up operations, mostly animations left to do

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{ulem}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\sl i.e.},~}
\newcommand{\eg}{{\sl e.g.},~}
\newcommand{\parentd}{../}
\newcommand{\R}{\mathbb{R}}
\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
\providecommand{\alert}[1]{\textbf{#1}}
\usetheme{Novartis}
\doNotUseNovartisLogoOnAllSlides
\doNotUseNovartisLogoOnTitlePage
\setbeamertemplate{blocks}[rounded][shadow=true]

\title{Matroids \& testing in families of dependent hypotheses}
\author{Florian Klinglmueller, Willi Maurer}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}
\frametitle{Matroids}

\maketitle

\begin{frame}
\frametitle{Outline}
\setcounter{tocdepth}{3}
\tableofcontents
\end{frame}

\section{Algebraically dependent hypotheses}

\begin{frame}
\frametitle{Some history:}
\framesubtitle{Testing hypotheses of no differences in pairwise
  comparisons of 3 treatments}

% Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):

\begin{block}{$H_1:\mu_1=\mu_2,\; H_2:\mu_2=\mu_3,\;H_3:\mu_1=\mu_3$}
  \begin{enumerate}
  \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
    \mu_1=\mu_2=\mu_3$ at level $\alpha$,
  \item test each $H_1,H_2,H_3$ at level $\alpha$,
  \end{enumerate}
  protects the FWER at level $\alpha$.
\end{block}
\pause
  \begin{itemize}
    \item  essentially closed test, since %; no need to test pairwise
                                %intersection hypotheses since 
        $H_{\{i,j\}} = H_{\{1,2,3\}}, i\neq j,$

    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
  \end{itemize}

\end{frame}

\begin{frame}
\frametitle{Some history}
\framesubtitle{Testing hypotheses of no difference in pairwise
    comparisons of $m>3$ treatments:}

  Modified Sequentially Rejective
  Multiple test procedures \cite{shaffer1986modified} % Example: Modified Bonferroni-Holm
      \begin{itemize}
      \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
      \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test
        remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ =
        max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are
        false.
      \end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{tabular}{ll}

  \begin{minipage}{.5\textwidth}
    \input{full4}
  \end{minipage} 
  \begin{minipage}{.5\textwidth}
    \begin{block}{Graphical representation}
      \begin{itemize}
      \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq
        m*(m-1)/2$

      \item Undirected graph %$\mathscr{G} = (V,E)$:
        \begin{itemize}
        \item vertices correspond to means, 
          %\emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
        \item edges correspond to hypotheses, 
          % \emph{i.e.}, $E = \left\{e_1,...,e_n\} \right\}$, $e_\ell = \{\mu_i,\mu_j\}$, $i
          % \neq j$, and $e_\ell \neq e_k$ for $\ell \neq k$
        \end{itemize}
      \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
      \item $I$ induces a subgraph %$\mathscr{G}_I$ 
        on the edge set
        %$E_I=\{e_i: i \in I\}$
      \end{itemize} % ends low level
    \end{block}
  \end{minipage}

\end{tabular}
\end{frame}

% \begin{block}{Example: $m=4, n=6$ :}
%       \begin{itemize}
%       \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
%       \item Dynamic: $t^*_j \le t_j$, hence more powerful than
%         "static"
%       \end{itemize}
%   \end{block}

\begin{frame}
\frametitle{Example: Modified Bonferroni Holm \cite{shaffer1986modified}}
\begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}

    \input{shaffer}
\end{minipage} &
  \begin{minipage}{.6\textwidth}
    \begin{block}{$m=4, n=6, \only<2->{j=2}$}
      % \begin{itemize}
      % \item "Static": $t_j$ = max number of true hypotheses if $j$ are
      %   false.
      % \item "Dynamic": $t^*_j $ = max. number of true hypotheses if
      %   $H_{i_1},...,H_{i_j}$ are false.
%      \item Example:  \only<3->{$ j = 2$ :}
        \begin{description}
        \item<3->[Static:]  $H_1$ and $H_2$ are false
            $\Rightarrow$ $t_2 = 3$ (at most $H_4, H_5 $ and $H_6$ can be
            simultaneously true)
        \item<7->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
          $t^*_2 = 2 < t_2 =3$ (only $H_3$, $H_4$, or $H_2$, $H_5$ can be
            simultaneously true)
        \end{description}

 %     \end{itemize} % ends low level
    \end{block}
  \end{minipage}
\end{tabular}
\end{frame}

% \begin{frame}
% \frametitle{Pairwise mean comparisons (two-sided)}
% \begin{itemize}

% \item Consider $\mu_i, i \in \{1,...,m\}$

% \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

% \item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
% \begin{itemize}
% \item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
% \item edges correspond to hypotheses, \emph{i.e.}, $E =
%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
%      j$, and $e_l \neq e_k$ for $l \neq k$
% \end{itemize}
% \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
% \item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
%  \in I\}$
% \end{itemize} % ends low level
% \end{frame}

% \begin{frame}
% \frametitle{Example 1: 4 Groups}
% \input{full4ext}
% \end{frame}

% \begin{frame}
% \frametitle{Example 1: 4 Means}
% \input{full4}
% \end{frame}

\begin{frame}
\frametitle{General contrast tests (two-sided)}
\begin{itemize}

\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$

\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
 c_{ij} = 0$
\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
 with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
   I}$
\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$

\item Elementary hypotheses are represented by column vectors,
 intersection hypotheses by sub-matrices of the contrast matrix

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 2: Contrast test}

\begin{tabular}{ll}
 \begin{minipage}{.44\textwidth}
   \begin{block}{\footnotesize Pairwise comparisons of 4 means}
     {\footnotesize
       \begin{displaymath}
         \bs{C}^t = \left(
           \begin{array}{rrrr}
             -1 & 1 & 0 & 0 \\
             -1 & 0 & 1 & 0 \\
             -1 & 0 & 0 & 1 \\
             0 & -1 & 1 & 0 \\
             0 & -1 & 0 & 1 \\
             0 & 0 & -1 & 1
           \end{array}
         \right)
       \end{displaymath}}
   \end{block}
 \end{minipage} &
 \begin{minipage}{.56\textwidth}
\pause
   \begin{block}{\footnotesize Modified Williams Contrasts for Dose Response \cite{bretz2010multiple}}
      {\footnotesize
       \begin{displaymath}
         \bs{C}^t=\left(
           \begin{array}{rrrr}
             1 & 0 & 0 & -1\\
             1 & 0 &-.51 & -.49 \\
             1 & -.35 & -.33 & -.32\\
             .51 & .49 & -.51 & -.48\\
             .51 & .49 & 0 &  -1\\
             .35 & .33 & .32& -1
           \end{array}\right)
       \end{displaymath}}
   \end{block}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Dependent null hypotheses}

\begin{itemize}
\item $H_i$ \alert{depends} on $H_I$: $H_{I \cup i} = H_{I}$
\item induces a \alert{dependence relation} $i \alert{\sim_H} I$ on index sets $I \in \mathcal{P}(S)$
\item $I$ \alert{independent} iff for all $i
  \in I$ $i \nsim_H I\setminus i$ (\ie $H_{I} \subsetneq H_{I\setminus i}$)
\pause
\item For pairwise comparisons $I$ independent iff there are no cycles
  in the corresponding graph
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
 column rank
\end{itemize} % ends low level
\end{frame}

\section{Matroids}

\begin{frame}
\frametitle{Matroids}
\begin{itemize}

\item Matroids generalize the structures of graphs and vector spaces,
 \eg cycles, bases, independent sets
\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
 structure
\item \alert{Cryptomorphism} \ie many ways to define a matroid:
 \begin{itemize}
 \item From a graph perspective: Circuit axioms.
 \item From a vector space perspective: Independence axioms.
 \end{itemize}
%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Circuit axioms}

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Circuits of a matroid $M$ on $S$:}
       A collection $\mathscr{C}$ of subsets of $S$
       for which holds
     \begin{enumerate}
     \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
     \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$
       then $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
     \end{enumerate}
   \end{block}

 \end{minipage} &

 \begin{minipage}{.5\textwidth}

** TODO shorten animation

   \only<2->{\input{cycle}}
 \end{minipage} 
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Independence axioms}

\begin{tabular}{ll}
 \begin{minipage}{.5\textwidth}
   \begin{block}{Independent sets of a matroid $M$ on $S$}
     A a collection $\mathscr{I}$ of subsets of $S$ for which
     holds
     \begin{enumerate}
     \item $\emptyset \in \mathscr{I}$,
     \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
       \mathscr{I}$,
     \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in
       I \setminus J$ such that $J \cup i \in \mathscr{I}$,
     \end{enumerate}
   \end{block}
 \end{minipage} &

 \begin{minipage}{.5\textwidth}

** TODO animate independent sets

 \input{independent}
% \only<2>{\input{cycle}}
% \invisible<1>{\alert{\footnotesize minimally dependent set!}}
 \end{minipage} 
\end{tabular}

\end{frame}

\begin{frame}
  \frametitle{Matroid structure}

** TODO add some pictures detailing general structure

  \begin{itemize}
  \item \alert{Circuits} are minimally dependent sets (remove any element to
    get an independent set)
  \item \alert{Independent sets} are sets for which no subset of elements is a
    circuit
  \item Maximally independent sets are \alert{bases} (Base axioms)
  \item Cardinality of largest independent subset is the \alert{rank} of a set
    (Rank axioms)
  \item Maximally dependent sets are called \alert{flats} (Closure axioms)
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Dependence axioms}
** TODO animate dependence
\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \input{dependent}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
 Relation $\sim_M$ on elements and subsets of $S$,
 if for any $j \in S$, and sets $I,K \subseteq S$ holds
 \begin{enumerate}
 \item $i \sim_M I$ for all $i \in I$
 \item if $j \sim_M I$ and $j \nsim_M I\setminus \{i\}$
   then $i \sim \{j\} \cup (I \setminus \{i\})$
 \item if $j \sim_M I$ and for all $i \in I$, $i \sim_M K$ then $j \sim_M K$
 \end{enumerate}
 then $\sim_M$ defines a matroid on $S$.
 \end{minipage}
\end{tabular}

\end{frame}

\section{Matroids from algebraically dependent hypotheses}
\begin{frame}
 \frametitle{Dependent hypotheses}
 \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
   \begin{itemize}
   \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
   \item<2-> Does this mean a family of algebraically dependent
     hypotheses defines a matroid $M$, such that $\sim_H = \sim_M$?
   \end{itemize}
 \end{block}
\end{frame}

\begin{frame}
\frametitle{Matroids from algebraically dependent hypotheses}

 \begin{block}{Two-sided pariwise comparisons}
   For a family of two-sided pairwise mean comparisons $H_i$, $i \in
   S$, an elementary hypotheses $H_j$ depends on intersection
   hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
   the edge set $\{e_i:i \in I\}$ of the \alert{cycle matroid} defined by
   the corresponding graph.
 \end{block}

 \begin{block}{Two-sided linear contrast tests}
   For a family of two-sided general contrast tests $H_i$, $i \in S$,
   an elementary hypothesis $H_j$ depends on intersection hypotheses
   $H_I$ iff the contrast vector $\bs{c_j}$ is a linear combination of
   column vectors of contrast matrix $\bs{C}_I$ which corresponds to
   the \alert{vectorial matroid} defined by $\bs{C}$ on $\mathbb{R}^m$.
 \end{block}

\end{frame}

\begin{frame}
 \frametitle{Matroids from algebraically dependent hypotheses}
 {\Large What is it good for?\\
 \invisible<1>{Structure}}
\end{frame}

% \begin{frame}
% \frametitle{Maximally independent sets}
%  \begin{itemize}
%  \item Bases are maximally independent sets of $S$
%  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
%  \item Circuits are minimally dependent sets, \ie each subset of $C$
%    with rank $\rho C - 1$ is a maximally independent set of $C$
%  \end{itemize}
% \end{frame}

% \begin{frame}
% \frametitle{Maximally dependent sets}
% \begin{itemize}
% \item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
% \item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
% \item Hyperplanes are flats of rank $\rho S - 1$
% \item All flats are given by closing hyperplanes under intersection
% \item Circuits in $M^*$ are the complements of hyperplanes in $M$
% \end{itemize} % ends low level
% \end{frame}

% \begin{frame}{Examples}
% \framesubtitle<1>{Independent set}
% \framesubtitle<2>{Base}
% \framesubtitle<4>{Dependent set}
% \framesubtitle<3>{Cycle}
% \framesubtitle<5>{Closure}
% \only<1>{\input{independent}}
% \only<2>{\input{base}}
% \only<4>{\input{dependent}}
% \only<3>{\input{cycle}}
% \only<5>{\input{closure}}
% \end{frame}

\begin{frame}
\frametitle{Flats}
\begin{itemize}

\item Flats are maximally dependent sets \ie $j \nsim I \forall j \in
  S\setminus I$
\item The flats of a matroid are closed under intersection
\item $S$ (global null hypothesis) is a flat
\item Every single element of $S$ (elementary hypothesis) is a flat
\item For each $I \subseteq S$, the $H_I$ as a subset of the parameter
  space is identical to the intersection hypotheses defined by it
  closure $H_{\sigma I}$ (\ie flats correspond to weakly exhaustive
  sets \cite{bergmann1988improvements})
\item Define a stepwise procedure on the flats of $M$
\end{itemize} % ends low level
\end{frame}

\begin{frame}
 \frametitle{Stepwise test procedure}
 \begin{block}{Multiple testing procedure for algebraically related hypotheses}
   Assume that a family of null hypotheses $\mathscr{H}$ with
   dependence relation $\sim_H$ defines a matroid $M$ on $S$. Then a
   test procedure that rejects elementary hypotheses $H_i$ if for all
   flats $E \subseteq S$ with $i \in E$, $H_E$ is rejected using a
   local level-$\alpha$ test, controls the FWER in the strong sense.
 \end{block}

\end{frame}

\begin{frame}
\frametitle{Old wine in new bottles?}

\begin{itemize}
\item \cite{shaffer1986modified} represents pairwise comparisons
 by set partitions
\item \cite{bergmann1988improvements} define the concepts of an
 exhaustive sets
\item \cite{bernhard1991computergestuetzte} shows properties of
 exhaustive sets that are basically \alert{closure axioms} of a matroid
 and provides algorithms to check each intersection hypotheses if is
 a flat using rank function.
\item \cite{westfall1997multiple} generates flats using dependence
 arguments in vector spaces.
\item \cite{weichert2000robuste} suggest an MTP based on clique
 decomposition of the graph, cliques in the graph are flats
\item \cite{westfall2007multiple} provide a branch and bound algorithm
 to generate all flats. They note that for pairwise comparisons all
 exhaustive hypotheses are \alert{partition hypotheses}
\item \cite{brannath2010shortcuts} provide shortcuts for stepwise
  procedures on \eg the set of exhaustive index sets
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Added benefits}
 \begin{itemize}
 \item Many results about exhaustive sets follow directly from
   corresponding results in matroid theory.
 \item Efficient algorithms to find the flats of specific and general
   matroids are available from matroid theory.
 \item \alert{Scum principle} \cite{welsh2010matroid} everything
   important happens at the top \ie it is sufficient to find the
   \alert{hyperplanes} of the corresponding matroid.
 \item The structure of flats provides some intuition to procedures
   that operate on exhaustive sets. 
 \end{itemize}
\end{frame}

\begin{frame}
 \frametitle{Stepwise multiple testing procedure on the lattice of
   flats}

\begin{tabular}{ll}
 \begin{minipage}{.3\textwidth}
   \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
   \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
   \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
   \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
   {\tiny wikipedia:Partition\_of\_a\_Set}
 \end{minipage} &
 \begin{minipage}{.7\textwidth}
 \begin{itemize}
 \item All flats are given by intersections of hyperplanes
   $\rightarrow$ knowing all hyperplanes remaining flats are given by
   intersection of index sets
 \item Hyperplanes can by found by algorithms that enumerate the
   circuits of the dual matroid \cite{boros2003algorithms}
 \item The partial order of flats defined by inclusion defines a
   geometric lattice
 \item<2-> Testing only a subset of elementary hypotheses corresponds
   to intervals
 \item<3-> Rejection of elementary hypotheses removes intervals
 \end{itemize}

 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}{Matroid procedures for pairwise mean comparisons}
\begin{tabular}{ll}
 \begin{minipage}{.3\textwidth}
   \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
    {\tiny wikipedia:Bell\_number}
 \end{minipage} &
 \begin{minipage}{.7\textwidth}
 \begin{itemize}
 \item Two-sided comparisons of $m$ means
 \item The cycle matroid of the graph representation
   provides equivalent dependence structure
 \item The flats of a (simple) complete graph are
   isomporphic to the partitions of a set of size $m$
 \item Efficient algorithms to generate all exhaustive sets by
   enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new}
 \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
   51, 202, 876, 4139, ...
 \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\section{Graphical approaches for algebraically dependent and
  contextually related hypotheses}

\begin{frame}
\frametitle{Reminder: Graphical approaches}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
   \end{tikzpicture}

 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item Provides weights $w(I)_i$ for each index $i$ and each
     non-empty subset $I \subset S$
   \item Weights are used to define a closed test procedure of
     weighted intersection hypotheses tests
   \item We have $\sum_{i \in S} w(I)_{i} \leq 1$ (conservative),
   \item and $w(I)_{i} \leq w(J)_{i}$ for $J \subseteq I$ (consonance).
   \end{itemize}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
 \frametitle{Example: Hierarchical LSD}
\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \input{lsd2.tex}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item 3 Hypotheses:
     \begin{enumerate}
     \item Placebo vs. High Dose
     \item Placebo vs. Low Dose
     \item High Dose vs. Low Dose
     \end{enumerate}
   \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
   \item Test $H_1$ at level $\alpha$
   \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
     simultaneously
   \item Can we generalize this?
   \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
 \frametitle{Fixed sequence test of pairwise mean comparisons}

** TODO Picture of hierarchical test hereabouts

\begin{block}{Theorem}
 Assume that $M(\mathscr{H})$ neither contains loops nor
 parallels. Assume that hypotheses $H_1$ through $H_r$ have been
 rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
 to be tested at level $\alpha_r$ according to a prefixed order.  If
 for some $1 \leq s \leq t$ holds
 \begin{equation}
   \label{eq:cond.hierarchical}
   \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
 \end{equation}
 hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
 at full level $\alpha_ra$.
\end{block}

\end{frame}

\begin{frame}
 \frametitle{Example: Multiple doses}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[overlay]
     \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
   \end{tikzpicture}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
   \end{tikzpicture}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
 \frametitle{General weighting strategies}
 \framesubtitle{QVA Example}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
   \end{tikzpicture}

 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{tikzpicture}[remember picture,overlay]
     \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
 \end{tikzpicture}
 \end{minipage}
\end{tabular}
\end{frame}

\begin{frame}
\frametitle{General weighting strategies}

\begin{enumerate}
\item Pretest using consonant weighted closed test procedure (\eg
  graphical approach), let $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
\item Compute all matroid hyperplanes $\mathscr{L}$ for the matroid
  restricted to the remaining hypotheses $I$
\item $R \gets R \cup \{i: p_i \leq \min\{ w(J)_{i}\alpha: J \in
  \mathscr{L} \} \}$
\item $A \gets \{\bigcup J: \min\{p_j / w(J)_{j}: j \in J\} > \alpha,
J \in \mathscr{L} \}$
\item $I \gets I \setminus R$
\item Compute hyperplanes of next smaller rank: $\mathscr{L} \gets \{J\in \mathscr{L}: \rho J = \rho
 \mathscr{L} - 1\}$
\item Remove interval of rejected hypotheses: $\mathscr{L} \gets \{J \in \mathscr{L}: J \cap R = \emptyset\}$
\item Reduce to interval of not accepted: $\mathscr{L} \gets \{J \in \mathscr{L}: L \nsubseteq A\}$
\item Repeat until $\rho \mathscr{L} = 0$
\item Reject $R$ and $I \setminus A$
\end{enumerate}
\end{frame}
\section{Matriods from general families of hypotheses}

\begin{frame}
\frametitle{General Results}
\begin{itemize}
\item We give sufficient necessary conditions for a general family of
  hypotheses $\mathscr{H}$ to define a matroid
\item Systems with one sided hypotheses, and parallel hypotheses (\eg
 combined non-inferiority/superiority) do not give matroids with
 $\mathscr{L}$ all strictly exhaustive
\item It is an open question under which conditions the set of
  strictly exhaustive sets is given by the flats of a matroid
\end{itemize}

\end{frame}

\begin{frame}
 \frametitle{One sided hypotheses}
 \begin{block}{One-sided linear contrast hypotheses}
   A family of {\em one sided linear contrast hypotheses} is defined
   by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and
   vector $\bs{b} \in \mathbb{R}^n$ that define a  family of null
   hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu}
   \leq b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th
   column of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
 \end{block}
\end{frame}

\begin{frame}
 \frametitle{Hyperplane arrangement}
 \begin{itemize}
 \item Each hypothesis $H_i$ corresponds to a hyperplane $h_i =
   \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$
 \item Each hyperplane $h_i$ divides $\mathbb{R}^m$ into two regions
   \begin{itemize}
   \item $h_i^+ = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} <
     b_i\}$
   \item $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x}
     > b_i\}$
   \end{itemize}
 \item Defines a sign pattern
   \begin{equation}
     \label{eq:sign}
     \delta_i(\bs{x}) = \left\{
       \begin{array}{ll}
         + & \text{if } \bs{x} \in h_i^+\\
         0 & \text{if } \bs{x} \in h_i \\
         - & \text{if } \bs{x} \in h_i^-\\
       \end{array}\right.
   \end{equation}
 \item $\delta_i$ is $-$ if $H_i$ false at $x$, or $0$ or $+$
   otherwise. $(\delta_1,...,\delta_n)$ tells us to which combination
   of true and false null hypotheses some $\bs{x}$ belongs.

 \item Set of all sign patterns defined by $\mathscr{H}$:
   \begin{equation}
     \label{eq:faces}
     \mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
   \end{equation}
 \end{itemize}
\end{frame}

\begin{frame}
 \frametitle{Example}

\begin{tabular}{ll}
 \begin{minipage}{.4\textwidth}
 \includegraphics[height=1\textwidth]{spheresfukuda.pdf}
 \end{minipage} &
 \begin{minipage}{.6\textwidth}
   \begin{itemize}
   \item 4 contrast hypotheses about 3 means
   \item Cut the hyperplanes with the sphere in $\R^3$ (makes
     visualisation easier)
   \item Cells of the sphere arrangment correspond to sign patterns
     of $\mathscr{H}$
   \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
   \end{itemize}
 \end{minipage}
\end{tabular}

\end{frame}

\begin{frame}
 \frametitle{Stepwise multiple testing procedure for one-sided linear
   contrast hypotheses}

 Consider a family of general contrast hypotheses $\mathscr{H} =
 \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
 $\mathscr{F}(\mathscr{H}) = \{ \bs{\delta}(\bs{x}): \bs{x} \in
 \R^m\}$ denote the corresponding set of sign patterns. Consider a
 set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
 which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
 corresponding decision rule for each elementary hypotheses $H_i$ is
 given by
 \begin{equation}
   \label{eq:signtest}
   \psi_i = \min_{F: F_i \geq 0} \varphi_F,
 \end{equation}
 which controls the FWER at level $\alpha$ in the strong sense.

\end{frame}

\begin{frame}
 \frametitle{Computation of $\mathscr{F}$}
 The question remains how to compute $\mathcal{F}$. For this observe
that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
just the face lattice of the corresponding oriented
matroid. Consequently, $\mathcal{F}$ can be found using algorithms
for arrangement construction from polyhedral programming
\cite{avis1992pivoting,avis1996reverse,ferrez2001cuts} which are
polynomial in time and input size and compact in space!

\end{frame}

\begin{frame}
 \frametitle{Example}
 3 ordered means
\end{frame}

\section{Summary \& Outlook}

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
\item Efficient algorithms from the realms of combinatorics and algebraic geometry
\item {\em Scum principle} everything important happens at the top
\item Stepwise procedure for algebraically dependent and contextually
  related hypotheses (graphical approach)
\item Promising results also for one-sided hypotheses
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outlook}
\begin{itemize}
\item Inclusion of procedures for two-sided tests in \texttt{gMCP} 
\item Implementation of one-sided problems using polyhedral
  programming libraries
\item Potential applications:
  \begin{itemize}
  \item \texttt{MCPmod} on \sout{steroids} matroids
  \item Model selection
  \item Logic regression
  \end{itemize}
\item Tests for intersection hypotheses based on maximally independent
  sets 
\end{itemize}
\end{frame}

\bibliographystyle{alpha}
\bibliography{\parentd paper/cleanrefs}
\end{document}